Bricking up and overgrowing backdoors in Intrusion Detection Systems

### 2019-09-10 ex2/paper/paper.tex
Update paper and fix minor things in code and plotting script

%%
%% This is file `sample-sigconf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigconf')
%%
%% IMPORTANT NOTICE:
%%
%% For the copyright see the source file.
%%
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf.tex.
%%
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%%
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[10pt,sigconf,letterpaper,dvipsnames]{acmart}
\acmYear{2019}
\copyrightyear{2019}
\acmConference{CoNEXT '19}{December 9-12, 2019}{Orlando, Florida, USA}

@@

\subsubsection{\gls{rf} Pruning}
\todo{Max: Redundant?} For intrusion detection, \gls{dt} based models or \glspl{rf} of \glspl{dt}  usually achieve higher accuracy than \glspl{mlp} \cite{meghdouri_analysis_2018} and moreover they are generally faster to train and possibly more interpretable. For \glspl{ids} we thus consider the removal of backdoors from \gls{dt} based models to be a more pressing issue than from \glspl{mlp}.

We thus propose a defense approach that can be used for similar use cases like the mechanisms recently proposed to remove backdoors from \glspl{cnn}. The concept behind this pruning defense is that leaves that are never used by samples in the validation dataset might be used by backdoored inputs. If these ``useless'' leaves are removed, performance of the classifier on the validation dataset should not decrease
%by a lot
significantly
while decisions that are used for the backdoor are hopefully removed. We develop several variants of our pruning defense with increasing levels of sophistication. These variants are pruning \todo{Remove space in the enumeration}
\begin{enumerate}
\item leaves based on their usage: This means that the least used leaf is pruned first and the most used one last.
\item only ``harmless'' leaves based on their usage: In the case of \glspl{ids} we assume that attackers want to make attacks appear to be benign.
\item only ``harmless'' leaves based on their usage and based on their depth if two leaves have an equal number of usages: The rationale behind this is that we assume that hand-crafted backdoors require fewer rules for a \gls{dt} to classify than regular samples. Thus, samples higher up the tree (with lower depth) are pruned first, if they have the same number of usages like another leaf with a higher depth.
\end{enumerate}
\begin{figure*}[h]
\includegraphics[width=\textwidth]{pruning_example.pdf}
\caption{Example of a \gls{dt} being pruned. To the left of each leaf is \textcolor{MidnightBlue}{the number of times a leaf was used} by validation samples and its \textcolor{Fuchsia}{depth} in the \gls{dt} is to the right of each leaf. It is also indicated that a sample is \protect\sethlcolor{green} \hl{harmless} (leaves with a 0 inside) because these are the ones our algorithm considers as candidates for pruning. The leaf that is going to be cut in each step has a dashed connection to its parent. After step 2 the algorithm stops as pruning the last remaining leaf would make the \gls{dt} non-functional.}
\label{fig:pruningExample}
\end{figure*}

\autoref{fig:pruningExample} shows an example of variant (3) of pruning applied to a tree
\begin{figure}[h]
\includegraphics[width=\columnwidth]{../prune_CAIA_backdoor_15/prune.pdf}
\caption{Performance metrics throughout the pruning process for \unsw{}. }
\label{fig:performancePruningRF}
\end{figure}
%\todo{Max: Which pruning algorithm actually?}
and \autoref{fig:performancePruningRF} shows the resulting accuracy for pruning according to variant (3) for the \unsw{} dataset. With sufficient neurons being pruned, the accuracy of the backdoor approaches zero. The accuracy for the regular data does not decrease; it even increases by
%thanks to the the pruning also
reducing overfitting. With a significantly smaller validation set, pruning takes longer to remove the backdoor
%until it works
but it still works. This is an important finding as a large validation dataset makes it hard for users without access to expensive hardware to perform the pruning.

@@

\bibliographystyle{ACM-Reference-Format}
\bibliography{bibliography}

>>>>

%%
%% This is file `sample-sigconf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigconf')
%%
%% IMPORTANT NOTICE:
%%
%%% For the copyright see the source file.
%%
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf.tex.
%%
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%%
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[10pt,sigconf,letterpaper,dvipsnames]{acmart}
\acmYear{2019}
\copyrightyear{2019}
\acmConference{CoNEXT '19}{December 9-12, 2019}{Orlando, Florida, USA}

@@

\subsubsection{\gls{rf} Pruning}
%\todo{Max: Redundant?} For intrusion detection, \gls{dt} based models or \glspl{rf} of \glspl{dt}  usually achieve higher accuracy than \glspl{mlp} \cite{meghdouri_analysis_2018} and moreover they are generally faster to train and possibly more interpretable. For \glspl{ids} we thus consider the removal of backdoors from \gls{dt} based models to be a more pressing issue than from \glspl{mlp}.

We propose a defense approach that can be used for similar use cases like the mechanisms recently proposed to remove backdoors from \glspl{cnn}. The concept behind this pruning defense is that leaves that are never used by samples in the validation dataset might be used by backdoored inputs. If these ``useless'' leaves are removed, performance of the classifier on the validation dataset should not decrease
%by a lot
significantly
while decisions that are used for the backdoor are hopefully removed. We develop several variants of our pruning defense with increasing levels of sophistication. These variants are pruning \todo{Remove space in the enumeration}
\begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
\item leaves based on their usage: This means that the least used leaf is pruned first and the most used one last.
\item only ``harmless'' leaves based on their usage: In the case of \glspl{ids} we assume that attackers want to make attacks appear to be benign.
\item only ``harmless'' leaves based on their usage and based on their depth if two leaves have an equal number of usages: The rationale behind this is that we assume that hand-crafted backdoors require fewer rules for a \gls{dt} to classify than regular samples. Thus, samples higher up the tree (with lower depth) are pruned first, if they have the same number of usages like another leaf with a higher depth.
\end{enumerate}
\begin{figure*}[h]
\includegraphics[width=\textwidth]{pruning_example.pdf}
\caption{Example of a \gls{dt} being pruned. To the left of each leaf is \textcolor{MidnightBlue}{the number of times a leaf was used} by validation samples and its \textcolor{Fuchsia}{depth} in the \gls{dt} is to the right of each leaf. It is also indicated that a sample is \protect\sethlcolor{green} \hl{harmless} (leaves with a 0 inside). The leaf that is going to be cut in each step has a dashed connection to its parent.}
\label{fig:pruningExample}
\end{figure*}

\autoref{fig:pruningExample} shows an example of variant (3) of pruning applied to a tree
\begin{figure}[h]
\includegraphics[width=\columnwidth]{../prune_CAIA_backdoor_15/prune_oh_d.pdf}
\caption{Performance metrics throughout the pruning process for \unsw{}. Considering only harmless leaves and pruning leaves with smaller depth first. The longer the dashes, the higher the fraction of the validation set being used (from 0.01 until 1).}
\label{fig:performancePruningRF}
\end{figure}
%\todo{Max: Which pruning algorithm actually?}
and \autoref{fig:performancePruningRF} shows the resulting accuracy for pruning according to variant (3) for the \unsw{} dataset. With sufficient neurons being pruned, the accuracy of the backdoor approaches zero. The accuracy for the regular data does not decrease; it even increases by
%thanks to the the pruning also
reducing overfitting. %With a significantly smaller validation set, pruning takes longer to remove the backdoor
%until it works
%but it still works. This is an important finding as a large validation dataset makes it hard for users without access to expensive hardware to perform the pruning. 
Also with only a very small validation dataset, the pruning still works as expected. For \cic{} we get similar results, but the pruning does not remove the backdoor completely (\~30\% backdoor accuracy remains). We attribute this to the fact, that in this dataset there are also regular flows which have $\text{stddev}(\text{TTL}) > 0$. And thus backdoored flows cannot always be sharply distinguished from regular flows. We find that pruning only harmless leaves is very beneficial for \unsw{} but not for \cic{}. However, considering depth when making the decision which leaf to prune first leads to the backdoor being removed considerably earlier in the pruning process.

@@

\bibliographystyle{ACM-Reference-Format}
\bibliography{bibliography}
%\bibliographystyle{plain}
